TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-03-05 article.tex
paper TTT: un detalle de la redacci'on
paper TTT: avanzamos en los ejemplos

\usepackage{bm} % \bm{\alpha} bold greek symbol
\usepackage[makeroom]{cancel} % \cancel{} \bcancel{} etc
\usepackage{wrapfig} % \begin{wrapfigure} Pone figura al lado del texto
\usepackage{mdframed}
\usepackage{algorithm}
\usepackage{comment}
%\usepackage{quoting}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{csvsimple}
\usepackage{listings}
\renewcommand{\lstlistingname}{Code}% Listing -> Algorithm
\usepackage{hyperref}
\usepackage{todonotes}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}

@@

\en{As we observed that the agent $i$ won, we know from the causal model that the hidden difference of performances was positive.}

@@

@@

\en{A difference of one $\beta$ between two actual skills, $ s_i - s_j = \beta $, is equivalent to 76\% probability of winning.}

@@

\en{And the probability of a draw in the games (\texttt{p\_draw}) is usually initialized with the observed frequency of draws.}

@@

\en{The class \texttt{History} is used to compute the posteriors and evidence of a sequence of events.}
\es{La clase \texttt{History} se usa para computar los posteriors y la evidencia de secuencias de eventos.}
%
\en{In the first example, we instantiate the class \texttt{History} with three players (\texttt{\small "a", "b", "c"}) and three games.}
\es{En el primer ejemplo, inicializamos una la clase \texttt{History} con tres jugadores (\texttt{"a","b","c"}) y tres partidas.}
%
\en{In the first game \texttt{"a"} beats \texttt{"b"}, in the second game \texttt{"b"} beats \texttt{"c"}, and in the third game \texttt{"c"} beats \texttt{"a"}.}
\es{En la primera partida \texttt{"a"} le gana a \texttt{"b"}, en la segunda \texttt{"b"} le gana a \texttt{"c"} y en la tercera \texttt{"c"} le gana a \texttt{"a"}.}
%
\en{In brief, all agents win one game and lose the other.}
\es{En resumen, todos los agentes ganan una partida y pierden la otra.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:history, caption={\en{Initialization of a \texttt{History}'s instance with a three events sequence.}\es{Inicialización de una instancia de \texttt{History} con una secuencia de tres eventos}}, belowskip=-1.0 \baselineskip, aboveskip=0.1cm]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r},belowskip=-0.77 \baselineskip]
c1 = list(c("a"),c("b"))
c2 = list(c("b"),c("c"))
c3 = list(c("c"),c("a"))
composition = list(c1,c2,c3)
\end{lstlisting}
\end{paracol}
\begin{lstlisting}[backgroundcolor=\color{all}]
h = History(composition, gamma=0.0)
\end{lstlisting}
%
\en{where the variables \texttt{c1}, \texttt{c2}, and \texttt{c3} model the composition of the each game using the names of the agents (i.e. identifiers), the variable \texttt{composition} is a list containing the three events, and the parameter \texttt{gamma} specifies how much the skills change over time (where 0 means no change).}
\es{donde la variable \texttt{c1}, \texttt{c2} y \texttt{c3} modela la composici\'on de cada partida usando los nombres de los agentes (i.e. identificadores), la variable \texttt{composition} es una lista que contiene las tres partidas, y el par\'ametro \texttt{gamma} especifica cuánto cambian las habilidad en el tiempo (donde 0 significa ningún cambio).}
%
\en{The results are implicitly defined by the order in which the game compositions are initialized: the teams appearing first in the list beat those appearing later.}
\es{El resultado queda definido implícitamente por el orden en el que la composiciones de las partidas fueron inicializadas: los equipos que aparecen primero en la lista vencen a los que aparecen después.}
%
\en{The rest of the parameters are initialized with the default values seen in code~\ref{lst:parameters}.}
\es{El resto de los parámetros se inicializa con los valores por defecto vistos en el código~\ref{lst:parameters}.}

@@

\en{After initialization, the class \texttt{History} immediately triggers the computation of TrueSkill estimates: it instantiates a new player for each name, a game for each event, and applies the dynamic factor between players' estimates.}

@@

\en{In this example, where all agents beat each other and their skills do not change over time, it would be reasonable for all estimates to be the same after observing the data.}

@@

\en{To access them we can call the method \texttt{learning\_curves()} of the class \texttt{History}, which returns a dictionary that is indexed by the names of the agents.}
\es{Para acceder acceder a ellas podemos llamar al m\'etodo \texttt{learning\_curves()} de la clase \texttt{HIstory}, que devuelve un diccionario indexado por los nombres de los agentes.}

@@

\en{Since each player wins one game and loses the other, a proper estimation process should return the same skill estimate for all agents, as no one is stronger than the others.}
\es{Como cada jugador gana una partida y pierde la otra, un proceso de estimación correcto debería devolver la misma estimación de habilidad para todos los agentes, ya que nadie es más fuerte que los demás.}
\es{A pesar de que lo razonable sería tener la misma estimaci\'on para todos los agentes, los posteriors reportados por TrueSkill difieren mucho entre jugadores y varían mucho para uno mismo jugador.}
\en{However, the final skills reported by TrueSkill differ greatly among players.}

@@

\en{In addition, the skill estimates of each player show strong variations between events.}

@@

\en{The first estimate of players \texttt{"a"} and \texttt{"b"} have the same uncertainty and the same absolute value as mean, being positive for the winning player and negative for the loser, because they contain only the information of the event in which \texttt{"a"} beats \texttt{"b"}.}

@@

\en{The second estimate of players \texttt{"a"} and \texttt{"b"} posteriors show a decrease in the uncertainty and the mean closer to zero, because they contain also the information of the events in which agent \texttt{"c"} is involved.}

@@

\en{Estimates are not equal as they should be in a case where no one is better than anyone else and where the order of the games should not matter since skills do not change over time.}

@@

\en{The TrueSkill Through Time model corrects this problem by propagating information throughout the system.}

@@

\en{To compute those estimates, we call the method \texttt{convergence()} of the class \texttt{History}.}

@@

\en{It can be seen that the TrueSkill Through Time estimates are not only the same between players, but also have less uncertainty.}

@@

\en{Note also that the estimates do not change over time following the absence of a dynamic factor, specified at the beginning by the null value of \texttt{gamma}.}
\es{Notar también que las estimaciones de los jugadores no cambian en el tiempo de acuerdo con la ausencia de factor dinámico, expresado al principio mediante el valor nulo de \texttt{gamma}.}

@@

\en{We now analyze a complex example in which a new player joins a population of already known players.}
\es{Ahora analizamos un ejemplo más complejo en el que un jugador nuevo se integra a una población de jugadores ya conocidos.}
%
\en{Suppose we know the true skill of all players.}
\es{Supongamos que nosotros conocemos la verdadera habilidad de todos los jugadores.}
%
\en{In this example we want to see how close the estimate is to the true skill.}
\es{En este ejemplo queremos ver cuan cercana es la estimación de la verdadera habilidad.}
%
\en{For this purpose, we establish a skill that changes over time following a logistic function, to be approximated by the package.}
\es{Para eso establecemos una habilidad que cambia en el tiempo siguiendo una función logísitica, que deberá ser aproximada por el paquete.}
%
\en{In the following code we generate the agent's learning curve and 1000 opponents' skills around the agent's skill (i.e. the agent plays with opponents of his own level).}
\es{En el siguiente código generamos la curva de aprendizaje del agente y 1000 oponentes con habilidades alrededor de la habilidad del agente (i.e. el agente juega con oponentes de su propio nivel).}
%
\begin{lstlisting}[backgroundcolor=\color{python}, label=lst:synthetic_skill, caption={\en{Synthetic learning curve}\es{Curva de aprendizaje sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
import math; from numpy.random import normal, seed; seed(99)
def skill(experience, middle, maximum, slope):
    return maximum/(1+math.exp(slope*(-experience+middle)))
agent = [skill(i, 500, 2, 0.0075) for i in range(1000)]
other = normal(agent,scale=0.5,size=1000)
\end{lstlisting}
%
\en{Here we only show the \proglang{Python} version because the amount of code is bigger than before (\proglang{Julia} and \proglang{R} versions can be found in the appendix \ref{AGREGA-CODE}).}
\es{Aquí mostramos solo la versión de \proglang{Python}, debido a que la cantidad de código es mayor que antes (las versiones de \proglang{Julia} y \proglang{R} se encuentran en el anexo \ref{AGREGA-CODIGO}).}
%
\en{The list \texttt{agent} represents the agent's skills, has a smooth growth starting at zero and ending at two, passing through one at event 500.}
\es{La lista \texttt{agent} representa las habilidades del agente, tiene un crecimiento suave que comienza en cero y termina en dos, pasando por uno en el evento 500.}
%
\en{The list \texttt{other} represents the randomly generated opponents' skills with a normal distribution, centered on the agent's skill with a standard deviation of $0.5$.}
\es{La lista \texttt{other} representa las habilidades de los oponentes generadas de forma aleatorias con una distribución normal, centrada en la habilidad del agente con un desvío estandar de $0.5$.}
%
\begin{lstlisting}[backgroundcolor=\color{python},label=lst:synthetic_parameters, caption={\en{Synthetic history}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
composition = [[["a"], [str(i)]] for i in range(1000)]
results = [[1,0] if normal(agent[i]) > normal(other[i]) else [0,1] for i in range(1000)]
times = [(i for i in range(1000)]
priors = dict([(str(i), Player(Gaussian(other[i], 0.2))) for i in range(1000)])

h = History(composition, results, times, priors, gamma=0.015)
h.convergence()
mu = [tp[1].mu for tp in h.learning_curves()["a"]] 
\end{lstlisting}
%
\en{In this code we define four variables to instantiate the class \texttt{History} in order to obtain the agent's learning curve.}
\es{En este código definimos cuatro variables para instancir la clase \texttt{History} con el objetivo de obtener la curva de aprendizaje del agente.}
%
\en{The variable \texttt{composition} contains 1000 games between the agent and 1000 different opponents.}
\es{La variable \texttt{composition} contiene 1000 partidas entre el agente y 1000 oponentes distintos.}
%
\en{The variable \texttt{results} contains the results randomly generated from the true skills: the player with the highest performance wins.}
\es{La variable \texttt{resultados} contiene los resultados generados aleatoriamente a partir de las habilidades reales: gana el jugador con mayor rendimiento.}
%
\en{The variable \texttt{time} contains integer values from 0 to 999, indicating the time period of each event.}
\es{La variable \texttt{time} contiene valores enteros de 0 a 999, indicando el período temporal de cada evento.}
%
\en{When we pass it to the class \texttt{History}, the difference between events determines the number of times the dynamic uncertainty ($\gamma^2$) should be applied as we go forward in time.}
\es{Cuando se la pasamos a la clase \texttt{History}, la diferencia entre partidas determina la cantidad de veces que la incertidumbre dinámica ($\gamma^2$) debe ser aplicada a medida que avanzamos en el tiempo.}
%
\en{La variable \texttt{priors} es un diccionario que permite personalizar las caracterísiticas de los jugadores.}
\es{La variable \texttt{priors} es un diccionario indexado por los nombres con priors de los oponetes (todas las estimaciones con baja incertidumbre).}
%
\en{In this example, we assign low uncertainty to the priors of the opponents to simulate that they are already known players.}
\es{En este ejemplo, le asignamos baja incertidumbre a los priors de los oponentes para simular que son jugadores ya conocidos.}
%
\en{The class \texttt{History} initializes the rest of the players (in this case only the agent) with default values and a dynamic factor \texttt{gamma=0.015}.}
\es{La clase \texttt{History} inicializa al resto de los jugadores (en este caso sólo al agente) con los valores por defecto y un factor dinámico \texttt{gamma=0.015}.}
%
\en{Finally, we converge the class \texttt{History} to obtain the TrueSkill Through Time estimates and the agent's learning curve.}
\es{Finalmente, hacemos converger la clase \texttt{Historia} para obtener las estimaciones de TrueSkill Through Time y la curva de aprendizaje del agente.}
%
\en{In the following figure we can see the difference between the estimated learning curve and the actual learning curve.}
\es{En la siguiente figura podemos ver la diferencia entre la curva de aprendizaje estimada y la curva de aprendizaje real.}

@@

@@

\en{In this example we show how to solve the complete history of the Association of Tennis Professionals (ATP).}
\es{En este ejemplo mostramos cómo resolver la historia de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has 447 thousand games that range from the year 1915 up to 2020, in which 19 thousand different players participate.}
\es{La base de datos cuenta con 447 mil partidas que van desde 1915 hasta el 2020 en las que participan 19 mil jugadores.}
%
\en{The \texttt{csv} file contains both single and double matches, indicated by the column \texttt{double}.}
\es{El archivo \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{We have an identifier for each game (\texttt{match\_id}) and its round number (\texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identidicador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{It contains four columns indicating the players' identifiers and four columns indicating their names: for example \texttt{w2\_id} is the identifier of the second player of the winning team (second players have null values in single games) and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally we have the name of the tournament (\texttt{tour\_name}), its identified (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the ground (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because is the most perfomant (\proglang{Python} and \proglang{R} versions can be found in the appendix \ref{AGREGA-CODE}).}
\es{Aquí mostramos solo la versión de \proglang{Julia}, debido a que es la más eficiente (las versiones de \proglang{Python} y \proglang{R} se encuentran en el anexo \ref{AGREGA-CODIGO}).}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

>>>>

\usepackage{bm} % \bm{\alpha} bold greek symbol
\usepackage[makeroom]{cancel} % \cancel{} \bcancel{} etc
\usepackage{mdframed}
\usepackage{algorithm}
\usepackage{comment}
%\usepackage{quoting}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{csvsimple}
\usepackage{listings}
\renewcommand{\lstlistingname}{Code}% Listing -> Algorithm
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage[binary-units]{siunitx}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}

@@

\en{As we observed that agent $i$ won, we know from the causal model that the hidden difference of performances was positive.}

@@

\todo[inline]{Armar una tabla con los par\'ametros default}

@@

\en{A difference of one $\beta$ between two actual skills, $ s_i - s_j = \beta $, is equivalent to \SI{76}{\percent} probability of winning.}

@@

\en{And the probability of a draw (\texttt{p\_draw}) is usually initialized with the observed frequency of draws.}

@@

\en{The class \texttt{History} is used to compute the posteriors and evidence of a sequence of events.}
\es{La clase \texttt{History} se usa para computar los posteriors y la evidencia de secuencias de eventos.}
%
\en{In the first example, we instantiate the class \texttt{History} with three players (\texttt{\small "a", "b", "c"}) and three games.}
\es{En el primer ejemplo, inicializamos una la clase \texttt{History} con tres jugadores (\texttt{"a","b","c"}) y tres partidas.}
%
\en{In the first game \texttt{"a"} beats \texttt{"b"}, in the second game \texttt{"b"} beats \texttt{"c"}, and in the third game \texttt{"c"} beats \texttt{"a"}.}
\es{En la primera partida \texttt{"a"} le gana a \texttt{"b"}, en la segunda \texttt{"b"} le gana a \texttt{"c"} y en la tercera \texttt{"c"} le gana a \texttt{"a"}.}
%
\en{In brief, all agents win one game and lose the other.}
\es{En resumen, todos los agentes ganan una partida y pierden la otra.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:history, caption={\en{Initialization of a \texttt{History}'s instance with a three events sequence.}\es{Inicialización de una instancia de \texttt{History} con una secuencia de tres eventos}}, belowskip=-1.0 \baselineskip, aboveskip=0.1cm]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r},belowskip=-0.77 \baselineskip]
c1 = list(c("a"),c("b"))
c2 = list(c("b"),c("c"))
c3 = list(c("c"),c("a"))
composition = list(c1,c2,c3)
\end{lstlisting}
\end{paracol}
\begin{lstlisting}[backgroundcolor=\color{all}]
h = History(composition, gamma=0.0)
\end{lstlisting}
%
\en{where the variables \texttt{c1}, \texttt{c2}, and \texttt{c3} model the composition of each game using the names of the agents (i.e. identifiers), the variable \texttt{composition} is a list containing the three events, and the parameter \texttt{gamma} specifies how much the skills change over time (where 0 means no change).}
\es{donde la variable \texttt{c1}, \texttt{c2} y \texttt{c3} modela la composici\'on de cada partida usando los nombres de los agentes (i.e. identificadores), la variable \texttt{composition} es una lista que contiene las tres partidas, y el par\'ametro \texttt{gamma} especifica cuánto cambian las habilidad en el tiempo (donde 0 significa ningún cambio).}
%
\en{The results are implicitly defined by the order in which the game compositions are initialized: the teams appearing first in the list beat those appearing later.}
\es{El resultado queda definido implícitamente por el orden en el que la composiciones de las partidas fueron inicializadas: los equipos que aparecen primero en la lista vencen a los que aparecen después.}
%
\en{The rest of the parameters are initialized with the default values seen in code~\ref{lst:parameters}.}
\es{El resto de los parámetros se inicializa con los valores por defecto vistos en el código~\ref{lst:parameters}.}

@@

\en{After initialization, the class \texttt{History} immediately triggers the computation of TrueSkill estimates: it instantiates a new player for each name, and a game for each event.}

@@

\en{In this example, where all agents beat each other and their skills do not change over time (as the dynamic factor is set to zero), all the estimates must be the same after observing the data as none is more skilled than the rest.}

@@

\en{To access them we can call the method \texttt{learning\_curves()} of the class \texttt{History}, which returns a dictionary indexed by the names of the agents.}
\es{Para acceder acceder a ellas podemos llamar al m\'etodo \texttt{learning\_curves()} de la clase \texttt{History}, que devuelve un diccionario indexado por los nombres de los agentes.}

@@

\en{The skills reported by TrueSkill present strong variations between events for each player and also among players for each event.}

@@

@@

\en{The estimates obtained after the first game in which \texttt{"a"} beats \texttt{"b"} have the same uncertainty and mean absolute value, being positive for the winner and negative for the loser.}

@@

\en{The estimates computed after the event in which each of the agents plays with \texttt{"c"} show a decrease in the uncertainties and the means are closer to zero.}

@@

\en{In this example where no one is better than rest, TrueSkill obtains different skill estimates, but they must be the same for all the players, }

@@

\en{TrueSkill Through Time solves this issue by propagating information throughout the system.}

@@

\en{Our package can also compute the estimates using TrueSkill Through Time by calling \texttt{convergence()} method of the class \texttt{History}.}

@@

\en{TrueSkill Through Time computes estimates that are not only the same for all players but also have less uncertainty.}

@@

% \en{Note also that the estimates do not change over time following the absence of a dynamic factor, specified at the beginning by the null value of \texttt{gamma}.}
% \es{Notar también que las estimaciones de los jugadores no cambian en el tiempo de acuerdo con la ausencia de factor dinámico, expresado al principio mediante el valor nulo de \texttt{gamma}.}

@@

\en{We now analyze a complex example in which a new player joins a large community of already known players.}
\es{Ahora analizamos un ejemplo más complejo en el que un jugador nuevo se integra a una población de jugadores ya conocidos.}
%
\en{Suppose we know the true skill of all players.}
\es{Supongamos que nosotros conocemos la verdadera habilidad de todos los jugadores.}
%
\en{In this example, we focus on the estimation of an evolving skill.}
\es{En este ejemplo queremos ver cuan cercana es la estimación de la verdadera habilidad.}
%
\en{For this purpose, we establish the skill of the target player to change over time following a logistic function. The community is generated ensuring each opponent has similar skill as the target player following her evolution.}
\es{Para eso establecemos una habilidad que cambia en el tiempo siguiendo una función logística, que deberá ser aproximada por el paquete.}
%
\en{In the following code, we generate the target player's learning curve and 1000 random opponents.}
\es{En el siguiente código generamos la curva de aprendizaje del agente y 1000 oponentes con habilidades alrededor de la habilidad del agente (i.e. el agente juega con oponentes de su propio nivel).}
%
\begin{lstlisting}[backgroundcolor=\color{python}, label=lst:synthetic_skill, caption={\en{Synthetic learning curve}\es{Curva de aprendizaje sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
import math; from numpy.random import normal, seed; seed(99)
def skill(experience, middle, maximum, slope):
    return maximum/(1+math.exp(slope*(-experience+middle)))
target = [skill(i, 500, 2, 0.0075) for i in range(1000)]
opponents = normal(target,scale=0.5,size=1000)
\end{lstlisting}
%
\en{We only include the \proglang{Python} version as \proglang{Julia} and \proglang{R} versions can be found in the appendix~\ref{AGREGA-CODE}).}
\es{Aquí mostramos solo la versión de \proglang{Python}, debido a que la cantidad de código es mayor que antes (las versiones de \proglang{Julia} y \proglang{R} se encuentran en el anexo \ref{AGREGA-CODIGO}).}
%
\en{The list \texttt{target} has the agent's skills at each moment. The values start at zero grow smoothly until the target player's skill reaches two.}
\es{La lista \texttt{agent} representa las habilidades del agente, tiene un crecimiento suave que comienza en cero y termina en dos, pasando por uno en el evento 500.}
%
\en{The list \texttt{opponents} includes the randomly generated opponents' skills following a normal distribution centered on each of the target's skills and a standard deviation of $0.5$.}
\es{La lista \texttt{other} representa las habilidades de los oponentes generadas de forma aleatorias con una distribución normal, centrada en la habilidad del agente con un desvío estandar de $0.5$.}
%
\begin{lstlisting}[backgroundcolor=\color{python},label=lst:synthetic_parameters, caption={\en{Synthetic history}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
composition = [[["a"], [str(i)]] for i in range(1000)]
results = [[1,0] if normal(target[i]) > normal(opponents[i]) else [0,1] for i in range(1000)]
times = [(i for i in range(1000)]
priors = dict([(str(i), Player(Gaussian(opponents[i], 0.2))) for i in range(1000)])

h = History(composition, results, times, priors, gamma=0.015)
h.convergence()
mu = [tp[1].mu for tp in h.learning_curves()["a"]] 
\end{lstlisting}
%
\en{In this code we define four variables to instantiate the class \texttt{History} in order to compute the target's learning curve.}
\es{En este código definimos cuatro variables para instancir la clase \texttt{History} con el objetivo de obtener la curva de aprendizaje del agente.}
%
\en{The variable \texttt{composition} contains 1000 games between the target player and different opponents.}
\es{La variable \texttt{composition} contiene 1000 partidas entre el agente y 1000 oponentes distintos.}
%
\en{To generate the list \texttt{results} we sample the performance following a normal distribution of the skills: the player with the highest performance wins.}
\es{La variable \texttt{resultados} contiene los resultados generados aleatoriamente a partir de las habilidades reales: gana el jugador con mayor rendimiento.}
%
% Parrafo

\en{The variable \texttt{time} contain a list of integer values ranging from 0 to 999 representing the time for each game.}
\es{La variable \texttt{time} contiene valores enteros de 0 a 999, indicando el período temporal de cada evento.}
%
\en{The class \texttt{History} uses this list to determine if the dynamic uncertainty ($\gamma^2$) should be applied or not. In this example, as all the events have different time, $\gamma^2$ is applied to all of them as we move forward in time.}
\es{Cuando se la pasamos a la clase \texttt{History}, la diferencia entre partidas determina la cantidad de veces que la incertidumbre dinámica ($\gamma^2$) debe ser aplicada a medida que avanzamos en el tiempo.}
%
\en{The variable \texttt{priors} is a dictionary which can be used to customize the properties of players.}
\es{La variable \texttt{priors} es un diccionario indexado por los nombres con priors de los oponetes (todas las estimaciones con baja incertidumbre).}
%
\en{In this example, we assign low uncertainty to the priors of the opponents to simulate we know their skill beforehand.}
\es{En este ejemplo, le asignamos baja incertidumbre a los priors de los oponentes para simular que son jugadores ya conocidos.}
%
\en{The class \texttt{History} initializes the rest of the players (in this case only the target player) using the default values and a dynamic factor \texttt{gamma=0.015}.}
\es{La clase \texttt{History} inicializa al resto de los jugadores (en este caso sólo al agente) con los valores por defecto y un factor dinámico \texttt{gamma=0.015}.}
%
\en{Finally, we use the \texttt{convergence()} method to obtain the TrueSkill Through Time estimates and the target's learning curve.}
\es{Finalmente, hacemos converger la clase \texttt{Historia} para obtener las estimaciones de TrueSkill Through Time y la curva de aprendizaje del agente.}

% Parrafo
%
\en{Figure~\ref{fig:logistic} presents the evolution of the estimate of the target player's skill. The solid black line represents the true target's skill, while the dashed black line shows the estimate. The dark grey zone depicts the uncertainty of the skill estimate, while the light grey indicates two times this uncertainty.}
\es{En la siguiente figura podemos ver la diferencia entre la curva de aprendizaje estimada y la curva de aprendizaje real.}

@@

\todo[inline]{Aca corresponde discutir qu\'e resultado dio y hacer un cierre diciendo que TTT es bueno para seguir la evolucion de los jugadores}

@@

\en{This last example is based on real-life data of all Association of Tennis Professionals (ATP) registered games.}
\es{En este ejemplo mostramos cómo resolver la historia de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players.}
\es{La base de datos cuenta con 447 mil partidas que van desde 1915 hasta el 2020 en las que participan 19 mil jugadores.}
%
\en{We condensed the information in a \texttt{csv} file containing both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{El archivo \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it has more performance than (\proglang{Python} and \proglang{R} versions, which can be found in the appendix \ref{AGREGA-CODE}).}
\es{Aquí mostramos solo la versión de \proglang{Julia}, debido a que es la más eficiente (las versiones de \proglang{Python} y \proglang{R} se encuentran en el anexo \ref{AGREGA-CODIGO}).}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")
